---
title: "R Notebook"
output: html_notebook
---

Data Ambassadors' attempts
```{r}
setwd("~/OneDrive - Teradata/Datathon/Group 3 intro/Data for USB/Rdata")
getwd()

```

```{r}
load("~/OneDrive - Teradata/Datathon/Group 3 intro/Data for USB/Rdata/civil_society_selected.Rdata")
load("~/OneDrive - Teradata/Datathon/Group 3 intro/Data for USB/Rdata/world_bank_selected.Rdata")
load("~/OneDrive - Teradata/Datathon/Group 3 intro/Data for USB/Rdata/world_bank_selected_vars_metadata.Rdata")
load("~/OneDrive - Teradata/Datathon/Group 3 intro/Data for USB/Rdata/civil_society_selected_vars_metadata.Rdata")
```

```{r}
#load("/Users/mm186159/OneDrive - Teradata/Datathon/group3/civil_society_selected.Rdata")
```

Load libraries
```{r}
library(ggplot2)
library(dplyr)
#install.packages("googleVis")
library(googleVis)
```

Join world bank with civil society 
```{r}
data_joined <- dplyr::inner_join(world_bank_selected,civil_society_selected,
                                 by=c("YEAR" = "YEAR","COUNTRY" = "COUNTRY"))
```

Select countries

```{r}
CEE_countries <- c('Albania',
                   'Bosnia and Herzegovina',
                   'Bulgaria',
                   'Czech Republic',
                   'Croatia',
                   'Hungary',
                   'Kosovo',
                   'Macedonia',
                   'Montenegro',
                   #'Serbia and Montenegro',
                   'Montenegro',
                   'Moldova',
                   'Poland',
                   'Romania',
                   'Slovakia',
                   'Czechoslovakia',
                   'Slovenia',
                   'Serbia',
                   'Austria'
)
data_joined_CEE <- data_joined %>% filter(COUNTRY %in% CEE_countries)
```

Select bunch of variables

```{r}
# vars_dots <- c("COUNTRY","YEAR",
#                "FP.CPI.TOTL.ZG","NY.GDP.MKTP.KD.ZG","SP.POP.TOTL","SH.DYN.NCOM.ZS","NY.GDP.PCAP.KD",
#                       "SH.DYN.0514__POPULATIONESTIMATESANDPROJECTIONS",
#             "SH.DYN.MORT__POPULATIONESTIMATESANDPROJECTIONS",
#             "SP.DYN.LE00.IN__POPULATIONESTIMATESANDPROJECTIONS","SI.DST.FRST.20",
#                       "EN.ATM.PM25.MC.ZS__SUSTAINABLEDEVELOPMENTGOALS",
#             "SHARE.OF.EMPLOYED.IN.TOTAL.POPULATION",
#             "SH.UHC.CONS.ZS","NW.HCA.PC",
#             "MEMBERS_ANY_PCT_WVS","CPI_SCORE_CPI",
#             "VOTER_TURNOUT_PRESIDENTIAL",
#             "VOTER_TURNOUT_PARLIAMENTARY",
#             "Q13.2_CIVIL_SOCIETY_TRADITIONS_BTI",
#             "SI_DEMOCRACY_STATUS_1_BTI",
#             "SII_MARKET_ECONOMY_STATUS_1_BTI",
#             "POPULATION_RED_CROSS",
#             "V2X_CSPART_VDEM",
#             "V2XCS_CCSI_VDEM",
#             "V2DLENGAGE_VDEM",
#             "V2CSREPRSS_VDEM",
#             "V2CSCNSULT_VDEM",
#             "V2CSPRTCPT_VDEM",
#             "V2CSANTIMV_VDEM",
#             "V2MEBIAS_VDEM",
#             "V2PEHEALTH_VDEM")

vars_dots <- c("COUNTRY","YEAR",
               "v2xcs_ccsi_VDEM",
               "v2pehealth_VDEM",
               "NY.GDP.MKTP.KD.ZG",
               "NY.GDP.PCAP.KD",
               "SP.DYN.LE00.IN__Populationestimatesandprojections",#Life expectancy at birth, total (years)",
               "si_democracy_status_1_BTI",#"Democracy status Index (1-10)",
               "sii_market_economy_status_1_BTI",#Market economy status Index (1-10)",
               "voter_turnout_presidential",
               "voter_turnout_parliamentary",
               "CPI_score_CPI"
               )
vars_dots <- toupper(vars_dots)

vars_underscores <- gsub("\\.","\\_",vars_dots)
```
select variables
```{r}
data_joined_CEE_sel <- data_joined_CEE %>% select(setdiff(vars_underscores,c("MEMBERS_ANY_PCT_WVS",
                                                                             "POPULATION_RED_CROSS")
                                                          )
                                                  )
```

join metadata

```{r}
world_bank_selected_vars_metadata2 <- world_bank_selected_vars_metadata 
names(world_bank_selected_vars_metadata2) <- names(civil_society_selected_vars_metadata)

METADATA <- rbind(civil_society_selected_vars_metadata,world_bank_selected_vars_metadata2)
# variables selected
Metadata_sel <- METADATA %>% filter(toupper(VARIABLE) %in% vars_dots)
```
select variables

select some years
e.g. 2017

```{r}
data_joined_CEE_sel_years <- data_joined_CEE_sel %>% filter(YEAR %in% c(2017))
```

Impute missing values with mean
```{r}
library(data.table)
ads_full <- data.table(data_joined_CEE_sel_years)
impute_mean_na <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
classes <- ads_full[,lapply(.SD,class),.SDcols = colnames(ads_full)] 
classes <- classes[1,] %>% as.vector
ads_full <- ads_full[,lapply(.SD,impute_mean_na),.SDcols = colnames(classes)[which(classes == 'numeric')]]

data_nonNA <- cbind(COUNTRY = data_joined_CEE_sel_years$COUNTRY, 
                    data.frame(ads_full)
)
```

let's get rid of missing values
```{r}
#data_nonNA <- data_joined_CEE_sel_years %>% filter(complete.cases(data_joined_CEE_sel_years))
```

check it
```{r}
dim(data_joined_CEE_sel_years)
```

```{r}
dim(data_nonNA)
```


Prepare biplots
get rid of varibales with zero variance
```{r}
vars<-sapply(data_nonNA %>% select(-COUNTRY,-YEAR),var)
vars0 <- vars[vars==0]

pcadata <- data_nonNA[,!names(data_nonNA)%in% names(vars0)]
```

PCA Analysis

```{r}
pca8_scaled <- prcomp(x = pcadata %>% 
                        select(-COUNTRY,-YEAR),
                               center = TRUE,scale.=TRUE)


biplot(pca8_scaled)

```
Percentage of variance
```{r}
plot(pca8_scaled)
```


by variables
```{r}
Cumulative.variance.explained <- cumsum(round(pca8_scaled$sdev^2/sum(pca8_scaled$sdev^2),2))
Cumulative.variance.explained[2]
```

Let's do something better for first 2 principal components
```{r}
pred <- as.data.frame(predict(pca8_scaled,newdata = data_nonNA))
pred <- data.frame(Country = as.character(pcadata$COUNTRY),pred[,1:2])

```

#first 2 
```{r}
pred
```

Join variables
```{r}
Metadata_sel2 <- Metadata_sel
Metadata_sel2$var_name <- toupper(gsub("\\.","\\_",Metadata_sel2$VARIABLE))
```

```{r}
variables <- as.data.frame(pca8_scaled$rotation)[,1:2]
variables$Var <- row.names(variables)
variables_joined <- left_join(variables,
                               Metadata_sel2 %>% 
                                 select(var_name,DESCRIPTION),by=c("Var"="var_name"))

```


Plot it finally

```{r}
factor <- 10
ggplot(variables_joined) + geom_point(aes(PC1*factor,PC2*factor),colour="red") + geom_text(aes(PC1*factor,PC2*factor,label=substr(DESCRIPTION,1,40)),size=2.5)+
  geom_point(data=pred,aes(PC1,PC2),colour="blue")+ geom_text(data=pred,aes(PC1,PC2,label=Country),size=3.5)+
  labs(title = paste0("2-dimensional mapping of CEE countries \n and civil society related metrics \n ",
                      "explaining ",as.character(round(100*Cumulative.variance.explained[2])),"% of variance"),
        y = "Second most important dimension", x = "First most important dimension") + 
  theme(plot.title = element_text(hjust = 0.5))

  
```
```{r}
ggplot(data_joined_CEE_sel_years)+geom_point(aes(VOTER_TURNOUT_PARLIAMENTARY,VOTER_TURNOUT_PRESIDENTIAL))+
  geom_text(aes(VOTER_TURNOUT_PARLIAMENTARY,VOTER_TURNOUT_PRESIDENTIAL+0.01,
                label=COUNTRY),size=2.5)
```
```{r}
ggplot(data_joined_CEE_sel_years)+geom_point(aes(VOTER_TURNOUT_PARLIAMENTARY,SI_DEMOCRACY_STATUS_1_BTI))+
  geom_text(aes(VOTER_TURNOUT_PARLIAMENTARY,SI_DEMOCRACY_STATUS_1_BTI+0.01,
                label=COUNTRY),size=2.5)+ylab("Democracy")
```

```{r}
#cormat <- as.data.frame(round(cor(data_joined_CEE_sel_years[,3:12]),2))
#head(cormat)
```


Moving chart of 

```{r}
names(data_joined_CEE_sel)
```

```{r}
data.google <- data_joined_CEE_sel %>% select(COUNTRY,YEAR,V2XCS_CCSI_VDEM,NY_GDP_PCAP_KD,
                                              VOTER_TURNOUT_PRESIDENTIAL,VOTER_TURNOUT_PARLIAMENTARY,
                                              SI_DEMOCRACY_STATUS_1_BTI)
data.google <- data.google[complete.cases(data.google),]
names(data.google) <- c("Country","year","VDEM_civil_society", 
                        "GDP_per_capita_constant_2010_US$",
                        "VOTER_TURNOUT_PRESIDENTIAL","VOTER_TURNOUT_PARLIAMENTARY",
                        "Democracy_status")

```

```{r}
myStateSettings <-'{"xZoomedDataMin":1199145600000,"colorOption":"2",
"duration":{"timeUnit":"Y","multiplier":1},"yLambda":1,
"yAxisOption":"4","sizeOption":"_UNISIZE",
"iconKeySettings":[],"xLambda":1,"nonSelectedAlpha":0,
"xZoomedDataMax":1262304000000,"iconType":"LINE",
"dimensions":{"iconDimensions":["dim0"]},
"showTrails":false,"uniColorForNonSelected":false,
"xAxisOption":"_TIME","orderedByX":false,"playDuration":15000,
"xZoomedIn":false,"time":"2010","yZoomedDataMin":0,
"yZoomedIn":false,"orderedByY":false,"yZoomedDataMax":100}' 
```

```{r}
chart <- gvisMotionChart(data.google, idvar="Country", 
                        timevar="year", 
                        colorvar ="Democracy_status", 
                        xvar = "GDP_per_capita_constant_2010_US$",
                        yvar = "VDEM_civil_society", 
                        sizevar="VOTER_TURNOUT_PRESIDENTIAL",
                                             date.format = "%Y",
                     options=list(state=myStateSettings))
```

```{r}
plot(chart)
```

